\sysfull

### 2018-04-15 paper/intro.tex
by Sinan - removed
intro some modifications

Energy-harvesting devices create unique challenges because they operate {\em
intermittently} when energy is
available~\citep{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
buffers energy in a small storage capacitor~\citep{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
down and recharges to operate again later. As an example, recharge times may be
tens of seconds in radio frequency powered medical device~\cite[Fig.
3c]{nadeau_naturebio_2017}.  The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor's~\cite{capybara} and some devices fail and restart operating $\approx$10 to
$\approx$100 times per second~\citep{tan_infocom_2016,mementos,nvp}.

\begin{wrapfigure}{t!}{0.5\textwidth}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/graffle/intro-figure.pdf} By Sinan
    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits and buffering updates to non-volatile (NV) variables in volatile (V) memory.}
    \label{fig:coalesce}
\end{wrapfigure}

\textbf{Data Consistency in Intermittent Computing.}  Software in an
energy-harvesting system operates according to the {\em intermittent execution
model}~\citep{dino,lucia_snapl_2017}, which corresponds to the
operation-failure-restart cycles of an energy-harvesting device. In an
intermittent software execution, an {\em operating period} proceeds for an
arbitrary duration (dictated by energy availability) before being interrupted
by a {\em failure period}. In a failure period, a device loses the volatile
state in its registers, stack, SRAM, and retains the state of any non-volatile
memory, such as FRAM or Flash. While capturing periodic
checkpoints~\citep{mementos,quickrecall} and sleep
scheduling~\citep{dewdrop,hibernus,hibernusplusplus} help preserve execution
progress, failures can leave non-volatile state incorrectly, partially updated,
and may leave checkpointed volatile state inconsistent with non-volatile state.
These inconsistencies cause intermittent execution to deviate from
continuously-powered behavior, often leading to an unrecoverable
application failure~\citep{dino,edb}.

Prior work developed two main approaches to dealing with data inconsistency for
intermittently-powered devices: (i) \emph{software-based programming and
execution models}~\citep{dino,ratchet,chain,alpaca} and (ii)
\emph{hardware-based computer architecture
support}~\citep{hicks_isca_2017,idetic,nvp}.  Complex architectural changes are
expensive to design, verify, and manufacture.  New architectures are also
inapplicable to existing systems~\citep{hicks_isca_2017,nvp}. Software
approaches are simpler and applicable to existing devices today, but prior
software approaches have a number of limitations that are the focus of the work
in this paper.  A key limitation of prior software approaches that we address in this paper is the {\em inflexibility} of a program statically decomposed into tasks.

\textbf{Task Decomposition of Intermittent Programs.} {\em
Task-based} programming and execution models require a
programmer~\citep{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
statically decompose a program into a collection of {\em tasks}.  A task can
include arbitrary computation and these existing systems guarantee that each
tasks executes {\em atomically}, despite arbitrarily-timed power failures.  
%
The programmer explicitly expresses task-to-task control flow, or the compiler
converts existing program control-flow into task-to-task control flow to sequence 
the tasks that it defines.
%
Figure~\ref{fig:coalesce} (top) illustrates how a program's tasks execute and
shows how tasks can impose a run-time overhead. 
%
At each task transition, the system incurs an overhead to track and atomically
commit modifications to the non-volatile memory, to maintain consistency of
program state~\citep{chain,alpaca}.  
%
The more task transitions a program executes at run time, the more overhead.

A savvy programmer may thus create very large tasks in an attempt to minimize
overhead by minimizing the number of task transitions.  However, a very large
task may require more energy to complete than a device's fixed hardware energy
buffer can hold.  Such a large task fails to completely execute assuming only
buffered energy is available (i.e., assuming the worst case).  While some
environments may have sufficient harvestable energy to replenish a device's
capacitor during operation, prolonging the execution period and enabling the
task to complete, assuming cooperation from the physical environment to avoid
non-termination is a risky programming proposition.  To eliminate this risk,
existing systems require the programmer or compiler to decompose a  program
into small tasks, all of which complete using only buffered energy.  These
constraints on task sizing leave an unsatisfying dilemma: large, efficient
tasks that risk non-termination, or small tasks that are guaranteed to
complete, but incur a high task transition and commit overhead.  In this work,
we propose a third way, using a novel technique called {\em dynamic task
coalescing} that efficiently executes small tasks, avoiding unnecessary
overheads while avoiding the risk of non-termination.

>>>>

Energy-harvesting devices create unique challenges because they operate {\em
intermittently} when energy is
available~\citep{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
buffers energy in a small storage capacitor~\citep{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
down and recharges to operate again later---corresponds to {\em intermittent execution model}~\citep{dino,lucia_snapl_2017} composed of
operation-power failure-restart cycles. As an example, recharge times may be
tens of seconds in radio frequency powered medical device~\cite[Fig.
3c]{nadeau_naturebio_2017}.  The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor's~\cite{capybara} and some devices fail and restart operating $\approx$10 to
$\approx$100 times per second~\citep{tan_infocom_2016,mementos,nvp}.

\begin{wrapfigure}{t!}{0.5\textwidth}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/graffle/intro-figure.pdf}
    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits. On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
    \label{fig:coalesce}
\end{wrapfigure}

\textbf{Data Consistency in Intermittent Computing.}  
%Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\citep{dino,lucia_snapl_2017}, which corresponds to the operation-failure-restart cycles of an energy-harvesting device. In an intermittent software execution, an {\em operating period} proceeds for anarbitrary duration (dictated by energy availability) before being interruptedby a {\em failure period}. In a failure period, 
Upon power failures, a device loses the volatile
state in its registers, stack, SRAM, and retains the state of any non-volatile
memory, such as FRAM or Flash. While capturing periodic
checkpoints~\citep{mementos,quickrecall} and sleep
scheduling~\citep{dewdrop,hibernus,hibernusplusplus} help preserve execution
progress, failures can leave non-volatile state incorrectly, partially updated,
and may leave checkpointed volatile state inconsistent with non-volatile state.
These inconsistencies cause intermittent execution to deviate from
continuously-powered behavior, often leading to an unrecoverable
application failure~\citep{dino,edb}. Prior work developed two main approaches to dealing with data inconsistency for
intermittently-powered devices: (i) \emph{software-based programming and
execution models}~\citep{dino,ratchet,chain,alpaca} and (ii)
\emph{hardware-based computer architecture
support}~\citep{hicks_isca_2017,idetic,nvp}.  Complex architectural changes are
expensive to design, verify, and manufacture.  New architectures are also
inapplicable to existing systems~\citep{hicks_isca_2017,nvp}. Software
approaches are simpler and applicable to existing devices today, but prior
software approaches have a number of limitations that are the focus of the work
in this paper.  A key limitation of prior software approaches that we address in this paper is the {\em inflexibility} of a program statically decomposed into tasks.

\textbf{Task Decomposition of Intermittent Programs.} {\em
Task-based} programming and execution models require a
programmer~\citep{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
statically decompose a program into a collection of {\em tasks}.  A task can
include arbitrary computation that should be executed despite arbitrarily-timed power failures.
%and these existing systems guarantee that each
%tasks executes {\em atomically}, despite arbitrarily-timed power failures.  
%
The programmer explicitly expresses task-to-task control flow, or the compiler
converts existing program control-flow into task-to-task control flow to sequence 
the tasks that it defines.
%
Figure~\ref{fig:coalesce} (top) illustrates how a program's tasks execute and
shows how tasks can impose a run-time overhead. 
%
At each task transition, the system incurs an overhead to track and atomically
\emph{commit} modifications to the non-volatile memory, to maintain consistency of
program state~\citep{chain,alpaca}.  
%
The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.

A savvy programmer may thus create very large tasks in an attempt to minimize
overhead by minimizing the number of task transitions.  However, a very large
task may require more energy to complete than a device's fixed hardware energy
buffer can hold.  Such a large task fails to completely execute assuming only
buffered energy is available (i.e., assuming the worst case).  While some
environments may have sufficient harvestable energy to replenish a device's
capacitor during operation, prolonging the execution period and enabling the
task to complete, assuming cooperation from the physical environment to avoid
non-termination is a risky programming proposition.  To eliminate this risk,
existing systems require the programmer or compiler to decompose a  program
into small tasks, all of which complete using only buffered energy.  These
constraints on task sizing leave an unsatisfying dilemma: large, efficient
tasks that risk non-termination, or small tasks that are guaranteed to
complete, but incur a high task transition and commit overhead.  In this work,
we propose a third way, using a novel technique called {\em adaptive task-execution by dynamic task
coalescing/splitting} that efficiently executes small tasks, avoiding unnecessary overheads while avoiding the risk of non-termination.
